---
title: "Do Steam Discounts Increase Player Engagement?"
subtitle: "A Difference-in-Differences Analysis"
author: "Wanxing Dai"

format:
  revealjs:
    embed-resources: true
    theme: simple
    slide-number: true
    transition: fade
    include-in-header:
      text: |
        <style>
        .reveal h1, .reveal h2 { color: #1b4f9c; }
        .reveal strong { color: #1b4f9c; }
        .reveal section { font-size: 0.82em; }

        .big-number {
          font-size: 2em;
          font-weight: bold;
          color: #1b4f9c;
        }

        .small-text {
          font-size: 0.62em;
          line-height: 1.25;
        }

        .reveal section.thank-you-slide h2 {
          font-size: 6rem !important;
          text-align: center !important;
          margin-top: 18vh !important;
          color: #1b4f9c !important;
        }
        </style>

execute:
  echo: false
  warning: false
  message: false
---
```{r}
price_dir <- "PriceHistory"
player_dir <- "PlayerCountHistoryPart1"

discount_games <- read.csv("discount_games_appids.csv", stringsAsFactors = FALSE)
discount_games$appid <- as.character(discount_games$appid)

app_info <- read.csv("applicationInformation.csv", stringsAsFactors = FALSE, check.names = FALSE)
app_info$appid <- as.character(app_info$appid)

get_game_name <- function(ids) {
  out <- app_info$name[match(ids, app_info$appid)]
  out[is.na(out) | out == ""] <- ids[is.na(out) | out == ""]
  out
}

get_file_path <- function(folder, appid) {
  possible <- c(file.path(folder, paste0(appid, ".csv")), file.path(folder, appid))
  existing <- possible[file.exists(possible)]
  if (length(existing) == 0) return(NA)
  existing[1]
}

read_players <- function(appid) {
  f <- get_file_path(player_dir, appid)
  if (is.na(f)) return(NULL)

  df <- read.csv(f, stringsAsFactors = FALSE, check.names = FALSE)

  if (!all(c("Time", "Playercount") %in% names(df))) return(NULL)

  out <- data.frame(
    appid = appid,
    date = as.Date(as.POSIXct(df$Time, format = "%Y-%m-%d %H:%M", tz = "UTC")),
    player_count = as.numeric(df$Playercount)
  )

  out <- out[!is.na(out$date) & !is.na(out$player_count), ]
  aggregate(player_count ~ appid + date, data = out, FUN = mean)
}

event_date <- as.Date("2019-06-25")
window_days <- 14
start_date <- event_date - window_days
end_date <- event_date + window_days

treated_ids <- unique(discount_games$appid[
  as.Date(discount_games$first_discount_date) == event_date
])

treated_candidates <- discount_games[discount_games$appid %in% treated_ids, ]
treated_candidates <- treated_candidates[order(-treated_candidates$max_discount), ]
treated_ids <- head(treated_candidates$appid, 20)

price_files <- list.files(price_dir, pattern = "\\.csv$", full.names = TRUE, recursive = TRUE)
player_files <- list.files(player_dir, pattern = "\\.csv$", full.names = TRUE, recursive = TRUE)

price_ids <- tools::file_path_sans_ext(basename(price_files))
player_ids <- tools::file_path_sans_ext(basename(player_files))

common_ids <- intersect(price_ids, player_ids)
control_pool <- setdiff(common_ids, discount_games$appid)

panel_treated <- data.frame()

for (appid in treated_ids) {
  players <- tryCatch(read_players(appid), error = function(e) NULL)
  if (is.null(players)) next

  players <- players[players$date >= start_date & players$date <= end_date, ]
  if (nrow(players) == 0) next

  players$treated <- 1
  players$relative_day <- as.numeric(players$date - event_date)
  players$post <- ifelse(players$date >= event_date, 1, 0)

  if (sum(players$post == 0) >= 7 & sum(players$post == 1) >= 7) {
    panel_treated <- rbind(panel_treated, players)
  }
}

control_panel <- data.frame()
max_controls_to_scan <- min(length(control_pool), 1000)

for (appid in control_pool[1:max_controls_to_scan]) {
  players <- tryCatch(read_players(appid), error = function(e) NULL)
  if (is.null(players)) next

  players <- players[players$date >= start_date & players$date <= end_date, ]
  if (nrow(players) == 0) next

  players$treated <- 0
  players$relative_day <- as.numeric(players$date - event_date)
  players$post <- ifelse(players$date >= event_date, 1, 0)

  if (sum(players$post == 0) >= 7 & sum(players$post == 1) >= 7) {
    control_panel <- rbind(control_panel, players)
  }
}

if (nrow(panel_treated) == 0) stop("No treated player data found.")
if (nrow(control_panel) == 0) stop("No control player data found.")

treated_meta <- aggregate(
  player_count ~ appid,
  data = panel_treated[panel_treated$post == 0, ],
  FUN = mean
)

control_meta <- aggregate(
  player_count ~ appid,
  data = control_panel[control_panel$post == 0, ],
  FUN = mean
)

names(treated_meta)[2] <- "pre_players"
names(control_meta)[2] <- "pre_players"

matched_pairs <- data.frame()
used_controls <- character(0)

for (i in seq_len(nrow(treated_meta))) {
  t <- treated_meta[i, ]
  available <- control_meta[!(control_meta$appid %in% used_controls), ]
  if (nrow(available) == 0) break

  distance <- abs(log1p(available$pre_players) - log1p(t$pre_players))
  best <- available[which.min(distance), ]

  matched_pairs <- rbind(
    matched_pairs,
    data.frame(
      pair_id = i,
      treated_appid = t$appid,
      control_appid = best$appid,
      treated_pre_players = t$pre_players,
      control_pre_players = best$pre_players,
      pre_difference = t$pre_players - best$pre_players
    )
  )

  used_controls <- c(used_controls, best$appid)
}

matched_pairs$treated_name <- get_game_name(matched_pairs$treated_appid)
matched_pairs$control_name <- get_game_name(matched_pairs$control_appid)

matched_ids <- c(matched_pairs$treated_appid, matched_pairs$control_appid)

panel_all <- rbind(panel_treated, control_panel)
panel_all <- panel_all[panel_all$appid %in% matched_ids, ]

panel_all$pair_id <- NA

for (i in seq_len(nrow(matched_pairs))) {
  panel_all$pair_id[panel_all$appid == matched_pairs$treated_appid[i]] <- matched_pairs$pair_id[i]
  panel_all$pair_id[panel_all$appid == matched_pairs$control_appid[i]] <- matched_pairs$pair_id[i]
}

panel_all <- panel_all[!is.na(panel_all$pair_id), ]
panel_all$log_players <- log1p(panel_all$player_count)

did_model <- lm(
  log_players ~ treated * post + factor(pair_id) + factor(relative_day),
  data = panel_all
)

did_result <- summary(did_model)

did_coef <- coef(did_result)["treated:post", "Estimate"]
did_p <- coef(did_result)["treated:post", "Pr(>|t|)"]
did_percent <- (exp(did_coef) - 1) * 100

raw_means <- aggregate(
  player_count ~ treated + post,
  data = panel_all,
  FUN = mean
)

raw_wide <- reshape(
  raw_means,
  idvar = "treated",
  timevar = "post",
  direction = "wide"
)

names(raw_wide) <- c("treated", "before", "after")
raw_wide$change <- raw_wide$after - raw_wide$before

raw_did <- raw_wide$change[raw_wide$treated == 1] -
  raw_wide$change[raw_wide$treated == 0]

control_before <- raw_wide$before[raw_wide$treated == 0]
control_after  <- raw_wide$after[raw_wide$treated == 0]
treat_before   <- raw_wide$before[raw_wide$treated == 1]
treat_after    <- raw_wide$after[raw_wide$treated == 1]

n_pairs <- length(unique(panel_all$pair_id))
n_days <- length(unique(panel_all$date))

n_treated_games <- length(unique(panel_all$appid[panel_all$treated == 1]))
n_control_games <- length(unique(panel_all$appid[panel_all$treated == 0]))

treated_game_list <- paste(matched_pairs$treated_name, collapse = "<br>")
control_game_list <- paste(matched_pairs$control_name, collapse = "<br>")
```

## Business Question

### Do Steam discounts increase player engagement?

Steam discounts are a major marketing tool.

The key question is:

**Do discounts actually increase player activity, or are discounted games already more popular?**

---

## Empirical Strategy

### Matched Difference-in-Differences

I compare discounted games with similar non-discounted games.

:::: {.columns}

::: {.column width="50%"}

### Treatment

Games discounted on **`r event_date`**

:::

::: {.column width="50%"}

### Control

Similar games not discounted during the same window

:::

::::

Games are matched based on pre-discount average player count.

---

## Matched Sample Check

:::: {.columns}

::: {.column width="25%"}

<div class="big-number">`r n_treated_games`</div>

Discounted games

:::

::: {.column width="25%"}

<div class="big-number">`r n_control_games`</div>

Control games

:::

::: {.column width="25%"}

<div class="big-number">`r n_pairs`</div>

Matched pairs

:::

::: {.column width="25%"}

<div class="big-number">`r n_days`</div>

Days

:::

::::

:::: {.columns}

::: {.column width="50%"}

### Discounted games

<div class="small-text">
`r treated_game_list`
</div>

:::

::: {.column width="50%"}

### Control games

<div class="small-text">
`r control_game_list`
</div>

:::

::::

---

## DID Design

$$
\log(Players_{it}) =
\beta_0 +
\beta_1 Discounted_i +
\beta_2 Post_t +
\beta_3(Discounted_i \times Post_t)
+ \epsilon_{it}
$$

### Key estimate

**$\beta_3$** measures the additional change in player engagement for discounted games after the discount event.

The model includes:

- Matched-pair fixed effects
- Relative-day fixed effects

---

## Parallel Trends Check

```{r, fig.width=9, fig.height=4.5}
trend <- aggregate(
  player_count ~ treated + relative_day,
  data = panel_all,
  FUN = mean
)

control_trend <- trend[trend$treated == 0, ]
treated_trend <- trend[trend$treated == 1, ]

ylim_range <- range(trend$player_count, na.rm = TRUE)

plot(
  control_trend$relative_day,
  control_trend$player_count,
  type = "l",
  lwd = 3,
  ylim = ylim_range,
  xlab = "Days Relative to Discount Event",
  ylab = "Average Player Count",
  main = "Player Count Trends Around Discount Event"
)

lines(
  treated_trend$relative_day,
  treated_trend$player_count,
  lwd = 3,
  lty = 2
)

abline(v = 0, lty = 3)

legend(
  "topleft",
  legend = c("Control", "Discounted"),
  lwd = 3,
  lty = c(1, 2),
  bty = "n"
)
```

---

## DID Result

```{r, fig.width=9, fig.height=4.5}
ylim_range <- range(
  c(control_before, control_after, treat_before, treat_after),
  na.rm = TRUE
)

plot(
  c(1, 2),
  c(control_before, control_after),
  type = "b",
  lwd = 3,
  pch = 19,
  ylim = ylim_range,
  xaxt = "n",
  xlab = "",
  ylab = "Average Player Count",
  main = "Before vs. After Discount Event"
)

lines(
  c(1, 2),
  c(treat_before, treat_after),
  type = "b",
  lwd = 3,
  lty = 2,
  pch = 19
)

axis(1, at = c(1, 2), labels = c("Before", "After"))

legend(
  "topleft",
  legend = c("Control", "Discounted"),
  lwd = 3,
  lty = c(1, 2),
  pch = 19,
  bty = "n"
)
```

### Estimated Effect

- Raw DID effect: **`r round(raw_did, 1)` players**
- Regression DID effect: **`r round(did_percent, 1)`%**
- p-value: **`r format.pval(did_p, digits = 3, eps = 0.001)`**

---

## Marketing Implications

### Recommendation 1

Use discounts as an engagement tool, not only a sales tool.

### Recommendation 2

Prioritize discounts for games with an existing player base that can be reactivated.

### Recommendation 3

Coordinate discounts with marketing campaigns, content updates, or community events to maximize engagement.

---

## Conclusion

### Key Takeaway

Steam discounts appear to meaningfully increase player engagement in the short run.

Matched DID helps separate the discount effect from pre-existing game popularity.

### Final point

Discounts should be evaluated not only by revenue impact, but also by their ability to reactivate players and increase platform engagement.

---

## Thank You {.thank-you-slide}